Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-29 thesis.tex
proof ch 6 inc new figures
Updates from Overleaf
finish proof reading
minor corrections / improvements
add extra gof plots for ATACCC
proof

% Thesis-specific maths commands
\newcommand{\dmax}{d_\text{max}}
\newcommand{\psens}{p_\text{sens}}
\newcommand{\ntot}{n_\text{tot}}
\newcommand{\ndet}{n_\text{d}}
\newcommand{\nnodet}{n_\text{u}}
\newcommand{\pnodet}{p_\text{u}}
\newcommand{\Npop}{N_\text{pop}}
\newcommand{\Ncis}{N_\text{CIS}}
\newcommand{\ncis}{\vec{n_\text{CIS}}}
\newcommand{\na}{\vec{n}_\text{a}}
\newcommand{\pcis}{\vec{p_\text{CIS}}}
\newcommand{\sched}{\mathbb{T}}
\newcommand{\nsched}{n_{\sched}}
\newcommand{\rate}{\mathfrak{R}}
\newcommand{\R}{\mathcal{R}_{0}}
\renewcommand{\Re}{\mathcal{R}_{e}(t)}
\newcommand{\Rc}{\mathcal{R}_{c}(t)}
\newcommand{\vlod}{v_\text{lod}}
\newcommand{\inform}{{_{\text{inform}}}}

@@

%% Bibliography
\usepackage[backend=biber,uniquename=init,sortcites=true,sorting=ynt,maxcitenames=2,style=apa,uniquelist=false]{biblatex}
\addbibresource{references.bib}
\addbibresource{refs-custom.bib}
% Display a linked DOI
\DeclareFieldFormat{doi}{%
  \mkbibacro{DOI}\addcolon\space
  \ifhyperref
    {\href{https://doi.org/#1}{\nolinkurl{#1}}}
    {\nolinkurl{#1}}%
}
% Declare the format for the pubstate field
\DeclareFieldFormat{pubstate}{\mkbibparens{#1}}
% Append pubstate to the end of the entry
\renewbibmacro*{addendum+pubstate}{%
  \printfield{addendum}%
  \newunit\newblock
  \printfield{pubstate}%
}

@@

    This relies on knowledge of the duration of positivity (how long individuals test positive for).
    Therefore, I first consider how to estimate the distribution of this duration in the general population.
    I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of positivity.
    % I take two complementary approaches using different datasets.
    % This model is fitted to data from a longitudinal study of contacts of individuals with confirmed SARS-CoV-2 infection.
    % In \cref{ATACCC}, I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of positivity.
    % This model is fitted to data from the ATACCC (Assessment of Transmission and Contagiousness of COVID-19 in Contacts) study, a longitudinal study of viral load in contacts of individuals with confirmed SARS-CoV-2 infection.
    I use this detailed model to exploit frequent (daily) sampling of individuals in a cohort of contact traced individuals.
    However, the cohort was small and the follow-up time short meaning that the tail of the duration distribution is poorly estimated.

@@

    The MRC Biostatistics Unit has been a great source of support.
    Many discussions, especially those with Drs Shaun Seaman and Brian Tom, contributed to progressing my work.
    % I would especially like to thank Drs Shaun Seaman and Brian Tom for their help with survival analysis.
    The admin team is a great asset to the Unit and were outstandingly supportive through the disrupted the pandemic caused.
    My fellow PhD students have created a great environment, having peers at the same stage in my journey has been invaluable.
    The Unit also arranged my funding, through the Bayes 4 Health grant, which I gratefully acknowledge.

@@

\subfile{MCMC-appendix}
\subfile{ATACCC-appendix-original-analysis}
\subfile{transmission-appendix-INLA}
\subfile{transmission-appendix-phase-type}
\subfile{transmission-appendix-plots}

>>>>

% Thesis-specific maths commands
\newcommand{\dmax}{d_\text{max}}
\newcommand{\psens}{p_\text{sens}}
\newcommand{\psenss}{p_\text{sens}^{(s)}}
\newcommand{\psensi}{p_\text{sens}^{(i)}}
\newcommand{\ntot}{n_\text{tot}}
\newcommand{\ndet}{n_\text{d}}
\newcommand{\nnodet}{n_\text{u}}
\newcommand{\pnodet}{p_\text{u}}
\newcommand{\Npop}{N_\text{pop}}
\newcommand{\Ncis}{N_\text{CIS}}
\newcommand{\ncis}{\vec{n_\text{CIS}}}
\newcommand{\na}{\vec{n}_\text{a}}
\newcommand{\pcis}{\vec{p_\text{CIS}}}
\newcommand{\sched}{\mathbb{T}}
\newcommand{\nsched}{n_{\sched}}
\newcommand{\rate}{\mathfrak{R}}
\newcommand{\R}{\mathcal{R}_{0}}
\renewcommand{\Re}{\mathcal{R}_{e}(t)}
\newcommand{\Rc}{\mathcal{R}_{c}(t)}
\newcommand{\vlod}{v_\text{lod}}
\newcommand{\inform}{{_{\text{inform}}}}

@@

%% Bibliography
\usepackage[backend=biber,uniquename=init,sortcites=true,sorting=ynt,maxcitenames=2,style=apa,uniquelist=false]{biblatex}
\setcounter{biburllcpenalty}{3000}
\addbibresource{references.bib}
\addbibresource{refs-custom.bib}
% Display a linked DOI
\DeclareFieldFormat{doi}{%
  \mkbibacro{DOI}\addcolon\space
  \ifhyperref
    {\href{https://doi.org/#1}{\nolinkurl{#1}}}
    {\nolinkurl{#1}}%
}
% Declare the format for the pubstate field
\DeclareFieldFormat{pubstate}{\mkbibparens{#1}}
% Append pubstate to the end of the entry
\renewbibmacro*{addendum+pubstate}{%
  \printfield{addendum}%
  \newunit\newblock
  \printfield{pubstate}%
}

@@

    This relies on knowledge of the duration of detectability (how long individuals test positive for).
    Therefore, I first consider how to estimate the distribution of this duration in the general population.
    I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of detectability.
    % I take two complementary approaches using different datasets.
    % This model is fitted to data from a longitudinal study of contacts of individuals with confirmed SARS-CoV-2 infection.
    % In \cref{ATACCC}, I extend a pre-existing model of how an individual's viral load changes over time to estimate the duration of detectability.
    % This model is fitted to data from the ATACCC (Assessment of Transmission and Contagiousness of COVID-19 in Contacts) study, a longitudinal study of viral load in contacts of individuals with confirmed SARS-CoV-2 infection.
    I use this detailed model to exploit frequent (daily) sampling of individuals in a cohort of contact traced individuals.
    However, the cohort was small and the follow-up time short meaning that the tail of the duration distribution is poorly estimated.

@@

    The MRC Biostatistics Unit has been a great source of support.
    Many discussions, especially those with Drs Shaun Seaman and Brian Tom, contributed to progressing my work.
    % I would especially like to thank Drs Shaun Seaman and Brian Tom for their help with survival analysis.
    The admin team is a great asset to the Unit and were outstandingly supportive through the disruption the pandemic caused.
    My fellow PhD students have created a great environment, having peers at the same stage in my journey has been invaluable.
    The Unit also arranged my funding, through the Bayes 4 Health grant, which I gratefully acknowledge.

@@

\subfile{MCMC-appendix}
\subfile{ATACCC-appendix-original-analysis}
\subfile{appendix-plots}
\subfile{cis-perf-new-ll}
\subfile{transmission-appendix-INLA}
\subfile{transmission-appendix-phase-type}

### 2024-03-29 transmission.tex
proof

$p_0$ increases at the same rate as individuals flow into $e_1$.
These individuals then transition into a set of compartments, $p_1, \dots, p_l$, each with their own transition rate.
I will choose $l$ and the transition rates such that combined length of stay (\ie time from entry in $p_1$ until exit from $p_l$) approximates the duration of positivity distribution estimated in \cref{E-imperf-test}.
This idea extends the technique of having multiple exposed or latent compartments (introduced in \cref{E-SEIR:sec:non-exponential}).
While the use of exponential or Erlang (gamma distributions with integer first parameter) is common in the observation model of compartmental models~\autocite[e.g.][]{overtonEpiBeds}, the use of more flexible phase-type distributions, as here, has rarely, if ever, been used in practice~\autocite{hurtadoGLCT}.
This broader class of distributions allows the model to capture the tail of the distribution, which is important when incidence suddenly decreases.

@@

Priors for the parameters are given in \cref{SEIR:table:priors}.
The parameters governing the duration of positivity were explained previously (see \cref{SEIR:table:ec-params}).

@@

An extension of either (or both) models into the rest of the epidemic would be useful.
The main obstacle is knowledge of how the duration of positivity distribution changes, and the need to incorporate the changes into the model.
The SEIR model would also need to include the effect of vaccination on transmission.

>>>>

$p_0$ increases at the same rate as individuals flow into $e_1$.
These individuals then transition into a set of compartments, $p_1, \dots, p_l$, each with their own transition rate.
I will choose $l$ and the transition rates such that combined length of stay (\ie time from entry in $p_1$ until exit from $p_l$) approximates the duration of detectability distribution estimated in \cref{E-imperf-test}.
This idea extends the technique of having multiple exposed or latent compartments (introduced in \cref{E-SEIR:sec:non-exponential}).
While the use of exponential or Erlang (gamma distributions with integer first parameter) is common in the observation model of compartmental models~\autocite[e.g.][]{overtonEpiBeds}, the use of more flexible phase-type distributions, as here, has rarely, if ever, been used in practice~\autocite{hurtadoGLCT}.
This broader class of distributions allows the model to capture the tail of the distribution, which is important when incidence suddenly decreases.

@@

Priors for the parameters are given in \cref{SEIR:table:priors}.
The parameters governing the duration of detectability were explained previously (see \cref{SEIR:table:ec-params}).

@@

An extension of either (or both) models into the rest of the epidemic would be useful.
The main obstacle is knowledge of how the duration of detectability distribution changes, and the need to incorporate the changes into the model.
The SEIR model would also need to include the effect of vaccination on transmission.
